### 2015-11-18 AssignmentVI/Assignment5-Ditz_Schroeder-WS15-16.tex
assignment 6 aufgabe 1a fertig

\begin{document}
%\header{Blatt}{Tutor}{Abgabedatum}{Vorlesung}{Bearbeiter}{Semester}{Anzahl Aufgaben}
\header{5}{Alexander Seitz}{16. November 2015}{Bioinformatics I}{\\Jonas Ditz \\\& Benjamin Schroeder}{WS 15/16}{4}

\section*{Theoretical Assignment - \textsl{Comparison with at most $l$ mismatches}}

General we are now only looking at the worst case scenario, which means uniform distributed mistmatches in an alignment. Also we only look at $k>=1$, as $k=0$ is in our application not a useful result. Assume two sequences of length t with l mismatches.\\
 Than both sequences contain $l$ tuples of length $L = \lfloor\frac{t}{l+1}\rfloor$ and one tuple which has a length of maximal $\lfloor\frac{t}{l+1}\rfloor$ . \\
\\
\noindent So both sequences share $l$ or $l+1$ k-tuples, if $k = \lfloor\frac{t}{l+1}\rfloor$, and for each 
$k \le \lfloor\frac{t}{l+1}\rfloor$ they share $(l+1) * \lfloor\frac{L}{k}\rfloor$
k-tuples.

\section*{Theoretical Assignment - \textsl{Linear programming by hand}}

The feasible region of this linear program is shown in figure \ref{feasibleRegion}, where the red line 
is constraint 1, green line is constraint 2, constraint 3 is drawn as a blue line and the yellow line 
is constraint 4.

\begin{figure}[ht]
 \centering
 \includegraphics[width=0.7\textwidth]{img/figure_1.png}
 \caption{feasible region of linear program (grey area)}
 \label{feasibleRegion}
\end{figure}

\subsection*{solution if t=s=1}
Generally we only used integers in this task. The constraint for the algorithm are as following if t and s value is 1 now:

\begin{eqnarray}
max. 1*x_1+1*x_2 \label{con1}\\ 
s.t. 2*x_1+x_2 \le 12 \label{con2}\\ 
2*x_1+3*x_2 \le 7 \label{con3}\\ 
-1*x_1+2*x_2 \le 16 \label{con4}\\ 
x_1,x_2 \ge 0 \label{con5}
\end{eqnarray}

\noindent the next step we started to compute each line of the tuple. Starting with constraint \ref{con3} under the aspect of constraint \ref{con5}. All possible tuples respectively to both constraints from the assignment are:\\
$temp1=\{(0,7);(1,6);(2,5);(3,4);(4,3);(5,2);(6,1);(7,0)\}$\\

\noindent the constraint \ref{con2} was applied to temp1:\\
$temp2=\{(0,7);(1,6);(2,5);(3,4);(4,3);(5,2);\}$\\

\noindent the next constraint (constraint \ref{con4}) was applied:\\
$temp3=\{(5,2)\}$\\

\noindent but not least the last constraint (4) is checked and the remaining tuple from the last step is valid:\\
$r=\{(5,2)\}$\\

\noindent the final result for the Algorithm with s=t=1 is the tuple: (5,2). Solving the formula with this numbers 
leads us to the following maximal value: $1*x_1 + 1*x_2 = 5+2 = 7$\\

\subsection*{How to make unsolvable?}
To get a not solvable algorithm for example the following constraint could be added:
	\begin{equation}\\
		x_1+x_2\ge 7
	\end{equation}

\noindent a new constraint just needs to be a contradiction to another constraint.

\subsection*{How to get infinite solutions with t and s?}
The linear program has infinite solutions, if we assign the following values to s and t:

%Getting infinite solutions,  requires to describe with the max function of a linear algorithm a line in the feasible region and not just one point.
%Therefore we decided to use constraint 3 and wanted to generate a max function along constraint (3). For the modelling of the max function we used two points from constraint (3) ($P_1(2,4); P_2(5,2)$) to determine $s=\frac{2}{3}$ and $t=1$ in the following equations.

%\begin{eqnarray*}
%	s * x_{1_{P_1}} + t*x_{2_{P_1}} = s * x_{1_{P_2}} + t*x_{2_{P_2}}\\
%\Leftrightarrow	s*2+t*4 = s*5+ t*2\\
%\Leftrightarrow	2t = 3s\\
%\Leftrightarrow	s = \frac{2}{3}t\\
%\rightarrow t=1\\	
%\end{eqnarray*}

%Now we controlled the resuld by inserting the extreme data points $P_1$ and $P_2$.

%\begin{eqnarray*}
%t=1\\
%\frac{2}{3}*t*x_1+t*x_2\\
%\Leftrightarrow\frac{2}{3}*x_1+x_2\\
%\\
%\frac{2}{3}*2+4 =\frac{2}{3}*5+2  = 5+\frac{1}{3}\\
%\end{eqnarray*}

%Not serious ;)
%The equation influence by t and s is:

\begin{align}
 s &= x_2 \nonumber \\
 t &= -x_1 \nonumber \\
\end{align}

\noindent These values would lead to the equation $x_2*x_1 + (-x_1)*x_2$, which is always zero. So 
the maximal result is zero and you can fill in arbitrary numbers to get this result, i.e. infinite 
solutions.

\subsection*{how to make the points a optimal solution?}

For the first point $P_1=(0,3)$ we have to set $s = -1$ and $t = 1$. That result in a maximal value 
for (0,3) because we cannot set $x_2$ arbitrary big due to constraint 4. And if we assign a bigger 
value to $x_1$, would decrease the result. Also we would violate constraint 2, if $x_1$ is to big.\\
\\
P2=(2,4)\\

\begin{figure}[h]
	\includegraphics[width=\textwidth]{img/figure-P24.png}
	\label{P24}
\end{figure}

To achive that P(2,4) is the only maximal value, we need to use the borders used in figure \ref{P24}. The lower one is described by the maximal constraint (4) (equation \ref{lowerbound})and the higher one is described by the horizontal line at $x2=4$. Next we can define relations, for the variables $m_2$ and $c_2$ of the searched linear equation. These are shown in equation \ref{rel1} and \ref{rel2}.

\begin{eqnarray}
-1x_1+2x_2=6\\
\Leftrightarrow  x_2 = \frac{1}{2} x_1 +3 \label{lowerbound}\\
=	x_2= m_1 x_1 + c_1\\
	\\
	m_1 > m_2 > 0 \label{rel1}\\
	c_1<c_2<4 \label{rel2}\\
\end{eqnarray}

 So any linear equation, which lies in the green area of figure \ref{P24} , is a possible solution. We will now just choose one and define $c_2 = 3.5 = \frac{7}{2}$ according to relation \label{rel2}. From this now we can deduce $m_2=\frac{1}{4}$ as shown in Equation \ref{m} so that P(2,4) is the only intersection with the feasible region.

\begin{eqnarray}
4=m_2 * 2+3.5\\
m_2 = \frac{4-3.5}{2}  = \frac{1}{4} \label{m}\\
\end{eqnarray}

The last step is now to get s and t, this is achieved by forming the found linear equation back to the desired form (Equation \ref{result24}). The complete result is for $s=\frac{-1}{14}$ and for $t= \frac{2}{7}$.

\begin{eqnarray}
x_2=\frac{1}{4} * x_1 + \frac{7}{2}\\
 	\Leftrightarrow 	- \frac{1}{4} x_1 + x_2=\frac{7}{2}\\
 	\Leftrightarrow 	- \frac{1}{14} x_1 + \frac{2}{7} x_2 = 1 \label{result24}\\
\Rightarrow s= - \frac{1}{14} ; t=\frac{2}{7}\\
\end{eqnarray}

\noindent P3=(1,3): We feel confident that there is no solution for this point.\\
\\
\noindent To make $P_4=(6,0)$ the maximum we set $s = 1$ and $t = -1$. We cannot set $x_1$ to a bigger value 
than 6 without violating constraint 1.

\section*{Theoretical Assignment - \textsl{Bonus: Carillo-Lipman bound}}

The Carrillo-Lipman bound is defined by $B_{pq} \coloneqq B(A) - (\sum_{i<j} s(\hat{A}_i, \hat{A}_j)-s(\hat{A}_p,\hat{A}_q))$, 
where $B(A)$ is a score for the alignment computed by a heuristic and $s(\hat{A}_i,\hat{A}_j)$ is 
the score of a optimal alignment of sequences $i$ and $j$. If we have just two sequences $p$ and $q$, 
this equation would be simplified as 
\[
B_{pq} = B(A) - (s(\hat{A}_p,\hat{A}_q)-s(\hat{A}_p,\hat{A}_q)) = B(A).
\]

\noindent So the Carrillo-Lipman bound for two sequences is just the computed score by a heuristic.
%\newpage
%\bibliography{Assignment4-Ditz_Schroeder-WS15-16.bib}
%\bibliographystyle{ieeetr}

>>>>

\begin{document}
%\header{Blatt}{Tutor}{Abgabedatum}{Vorlesung}{Bearbeiter}{Semester}{Anzahl Aufgaben}
\header{6}{Alexander Seitz}{23. November 2015}{Bioinformatics I}{\\Jonas Ditz \\\& Benjamin Schroeder}{WS 15/16}{2}

\section*{Theoretical Assignment - \textsl{Sequence-profile alignment and expected patterns in sequences}}

\subsection*{}
The profile as a PSWM for the given MSA would look like table \ref{tab:pswm}.

\begin{table}[h]
\centering
\caption{PSWM of the given MSA}
\begin{tabular}{c|ccccc}
   & $p_1$ & $p_2$ & $p_3$ & $p_4$ & $p_5$ \\
 \hline  
 \textbf{A} & $0.\overline{3}$ & 0 & $0.\overline{3}$ & 0 & 0 \\
 \textbf{C} & 0 & 0 & 0 & 1 & 0 \\
 \textbf{G} & 0 & 0 & 0 & 0 & 1 \\
 \textbf{T} & 0 & 1 & $0.\overline{6}$ & 0 & 0 \\
 \textbf{-} & $0.\overline{6}$ & 0 & 0 & 0 & 0 
 \label{tab:pswm}
\end{tabular}
\label{tab:pswm}
\end{table}

\noindent Using this PSWM we can now compute an optimal semiglobal alignment of our profile with the 
sequence $A = CATTCCGTTC$. First we calculate the scoring matrix using as a scoring function 
$s(a,b) = -1$, $s(a,a) = 3$ and $d = 2$:

\bigskip

\begin{table}[h]
\centering
\begin{tabular}{c|cccccccccc}
   & $b_1$ & $b_2$ & $b_3$ & $b_4$ & $b_5$ & $b_6$ & $b_7$ & $b_8$ & $b_9$ & $b_{10}$ \\
   & C & A & T & T & C & C & G & T & T & C \\
 \hline
 $p_1$ & $-1.\overline{6}$ & $-0.\overline{3}$ & $-1.\overline{6}$ & $-1.\overline{6}$ & $-1.\overline{6}$ & $-1.\overline{6}$ & $-1.\overline{6}$ & $-1.\overline{6}$ & $-1.\overline{6}$ & $-1.\overline{6}$ \\
 $p_2$ & $-1$ & $-1$ & $3$ & $3$ & $-1$ & $-1$ & $-1$ & $3$ & $3$ & $-1$ \\
 $p_3$ & $-1$ & $0.\overline{3}$ & $1.\overline{6}$ & $1.\overline{6}$ & $-1$ & $-1$ & $-1$ & $1.\overline{6}$ & $1.\overline{6}$ & $-1$ \\
 $p_4$ & $3$ & $-1$ & $-1$ & $-1$ & $3$ & $3$ & $-1$ & $-1$ & $-1$ & $3$ \\
 $p_5$ & $-1$ & $-1$ & $-1$ & $-1$ & $-1$ & $-1$ & $3$ & $-1$ & $-1$ & $-1$ 
\end{tabular}
\end{table}
\newpage
\noindent Now we fill the DP matrix using that scoring matrix:
\bigskip
\begin{table}[h!]
 \centering
 \begin{tabular}{c|ccccccccccc}
    & 0 & C & A & T & T & C & C & G & T & T & C \\
  \hline
  0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
  $p_1$ & -2 & $-1.\overline{6}$ & \textcolor{red}{$-0.\overline{3}$} & $-1.\overline{6}$ & $-1.\overline{6}$ & $-1.\overline{6}$ & $-1.\overline{6}$ & $-1.\overline{6}$ & $-1.\overline{6}$ & $-1.\overline{6}$ & $-1.\overline{6}$ \\
  $p_2$ & -4 & $-3$ & $-2.\overline{6}$ & \textcolor{red}{$2.\overline{6}$} & $1.\overline{3}$ & $-0.\overline{6}$ & $-2.\overline{6}$ & $-2.\overline{6}$ & $1.\overline{3}$ & $1.\overline{3}$ & $-0.\overline{6}$ \\
  $p_3$ & -6 & $-5$ & $-2.\overline{6}$ & $0.\overline{6}$ & \textcolor{red}{$4.\overline{3}$} & $2.\overline{3}$ & $0.\overline{3}$ & $-1.\overline{6}$ & $-0.\overline{6}$ & $3$ & $1$ \\
  $p_4$ & -8 & $-3$ & $-4.\overline{6}$ & $-1.\overline{3}$ & $2.\overline{3}$ & \textcolor{red}{$7.\overline{3}$} & \textcolor{red}{$5.\overline{3}$} & $3.\overline{3}$ & $1.\overline{3}$ & $1$ & $6$ \\
  $p_5$ & -10 & $-9$ & $-4$ & $-3.\overline{3}$ & $0.\overline{3}$ & $5.\overline{3}$ & $6.\overline{3}$ & \textcolor{red}{$8.\overline{3}$} & $6.\overline{3}$ & $4.\overline{3}$ &  $4$ \\
 \end{tabular} 
\end{table}

\noindent One can see that the optimal alignment (colored in red) is:

\begin{table}[h]
\centering
\begin{tabular}{cccccccccc}
 C&A&T&T&C&C&G&T&T&C\\
  &$p_1$&$p_2$&$p_3$&$p_4$&-&$p_5$& & & 
\end{tabular}
\end{table}

\section*{Theoretical Assignment - \textsl{Practice writing an introduction / background for a paper}}

%\newpage
%\bibliography{Assignment4-Ditz_Schroeder-WS15-16.bib}
%\bibliographystyle{ieeetr}
